Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes

### 2020-05-17 Ben_Manuscripts/hdphmm/Draft.tex
update draft.

  We applied the HDP-AR-HMM independently to all 24 trajectories of each of the four solutes studied.
  The number of states found for each trajectory varied between 5 and 30.  %BJC4: MET: (5, 28), need to finalized
  In Figure~\ref{fig:rz_unclustered}, we show the state sequence determined by the HDP-AR-HMM for
  an example methanol trajectory. The model distinguishes many of the states due to 
  differences in their mean levels while their fluctuations about those means are often 
  similar. In Section~\ref{section:mechanisms}, we use clustering to reduce the total state
  space for each solute in order to more easily interpret the state parameters.

@@

  As shown in Figure~\ref{fig:qualitative_unclustered}, realizations of our model 
  can produce trajectories that show qualitatively similar hopping and trapping 
  behavior to the MD trajectories to which they were fit. It is important that our
  model reproduces this behavior so that we can have confidence in its usefulness
  for quantitative predictions.

  % BJC4: This figure needs to be redone with new trajectory generation procedure
  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{qualitative_unclustered_MET2.pdf}
  \caption{We can use realizations of the HDP-AR-HMM to generate trajectories which 
  bear qualitative resemblance to the MD trajectories. We used the parameters 
  of the HDP-AR-HMM, whose state segments are visualized in 
  Figure~\ref{fig:rz_unclustered}, in order to construct an ensemble of 
  characteristic trajectories. Solute trajectories generated with
  the HDP-AR-HMM (blue) show the same hopping and trapping behavior exhibited by 
  MD (black).
  }\label{fig:qualitative_unclustered}
  \end{figure}

  In Figure~\ref{fig:unclustered_msds}, we show that in most cases we can reproduce the 
  MD MSDs on long timescales within uncertainty using realizations of the 
  HDP-AR-HMM.
  \begin{itemize}
    \item We generated the MSD curves using Method 1 described in 
    Section~\ref{method:realizations} of the methods.
    \item Our ability to reproduce these curves gives use confidence
    in their long term predictions.
    \item We underestimate the MSD of urea because of a small number of MD 
    trajectories with uncharacteristically high MSDs. Despite our best efforts, 
    we could not adequately parameterize an HDP-AR-HMM which reproduced said 
    behavior. (see SI)  %BJC4: TODO
    \item On long timescales, the model appear to have the same slope as 
    the MD trajectory.
  \end{itemize}

@@

@@

  % BJC4: Not sure if this is useful.
  By studying the parameters, it is now easier to describe the state behavior.
  \begin{enumerate}[label={State \theenumi :}, leftmargin=3.5\parindent]
     \item This state has intermediate dwell times with high past-hop dependence.
     The radial mean and evidence from Figure~\ref{fig:clustered_traj_MET} suggests
     that this state is common in the tail and pore region.
     \item This state seems to occur exclusively in the tail region based on its radial
     mean. Its fluctuations are small and there is little past hop dependence
     \item This is likely a hopping state that strongly contribute to the solute's MSD. 
     It is short lived and takes very large hops. This type of behavior is exhibited 
     near and far from the pore center as discussed earlier.
     \item This is a trapping state with a strong dependence on its previous fluctuation.
     Evidence from Figure~\ref{fig:clustered_traj_MET} and a radial mean of 1.8 suggest
     that it tends to occur at intermediate distances from the pore center.
     \item This is an intermediate-length trapping state that occurs close to the pore
     center. It's fluctuations are tight and relatively independent of its previous 
     fluctuation.
  \end{enumerate}

@@

  \textit{Relating Parameters to Solute-Membrane Interactions}: Based on the
  molecules and ions which constitute the membrane, it is straightforward to 
  hypothesize two types interactions that might occur.
  \begin{itemize}  
    \item All of the solutes are polar and have hydrogen bond donating and 
    accepting groups. 
    \item The monomers have 10 oxygen atoms distributed from head to tail that are 
    capable of accepting hydrogen bonds. 
    \item Also, the ionic bond between the head groups and sodium ions can
    be easily broken since both components can be stabilized by surrounding 
    water molecules and solutes.
    \item Therefore, it is logical to search for hydrogen bonding interactions
    and solute association with sodium ions.
  \end{itemize}

  For methanol, hydrogen bonding is a much more dominant interaction than sodium
  ion association. 
  \begin{itemize}
    \item In Figure~\ref{fig:hbond_fractions}, we show that methanol participates
    in hydrogen bonds to various parts of the monomer between 27 and 45\% of the
    total time it occupies the dominant states.
    \item Methanol associates with sodium ions less than 10\% of the time in 
    dominant states.
  \end{itemize}

@@

  The local number density of heavy atoms surrounding solutes help round out methanol's
  mechanistic picture.
  \begin{itemize}
    \item Although hydrogen bonding interactions are relatively frequent, methanol molecules
    are unbound by any measurable electrostatic interactions about half or more of the 
    simulation time.  
    \item Figure~\ref{fig:local_densities} helps to illustrate how the local density of 
    membrane components can aid in solute entrapment.
    \item When solutes experience long trapping times, the local number density of heavy 
    atoms is larger.
    \item Even when hydrogen bonds break, surrounding alkane density likely helps keep
    the solutes in place allowing them to eventually reform. This can happen deep in the
    tails or close to the monomer head groups.
    \item Methanol molecules which experience State 3 have by far the smallest local 
    number density which is likely responsible for their mobility, as suggested by the state's
    large covariance.
  \end{itemize}

>>>>

  We applied the HDP-AR-HMM independently to all 24 trajectories of each of the four solutes studied.
  The number of states found for each trajectory varied between 5 and 30.  %BJC4: MET: (5, 28), need to finalized
  Figure~\ref{fig:rz_unclustered}, shows the state sequence determined by the HDP-AR-HMM for
  an example methanol trajectory. The model distinguishes many of the states due to 
  differences in their mean levels while their fluctuations about those means are often 
  similar. In Section~\ref{section:mechanisms}, we use clustering to reduce the total state
  space for each solute in order to more easily interpret the state parameters.

@@

  As shown in Figure~\ref{fig:qualitative_unclustered}, realizations of our model 
  %MRS: would it be accurate to show ``produces trajectories that show . . . ``  Saying that it ``can 
  %produce trajectories'' is a little weasly - sounds like it usually doesn't.  
  % 
  can produce trajectories that show qualitatively similar hopping and trapping 
  behavior to the MD trajectories to which they were fit. It is important that our
  model reproduces this behavior so that we can have confidence in its usefulness
  for quantitative predictions.

  % BJC4: This figure needs to be redone with new trajectory generation procedure
  %MRS: is this still the old one?
  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{qualitative_unclustered_MET2.pdf}
  \caption{We can use realizations of the HDP-AR-HMM to generate trajectories which 
  bear qualitative resemblance to the MD trajectories. We used the parameters 
  of the HDP-AR-HMM, whose state segments are visualized in 
  Figure~\ref{fig:rz_unclustered}, in order to construct an ensemble of 
  characteristic trajectories. Solute trajectories generated with
  the HDP-AR-HMM (blue) show the same hopping and trapping behavior exhibited by 
  MD (black).
  }\label{fig:qualitative_unclustered}
  \end{figure}

  In Figure~\ref{fig:unclustered_msds}, we show that in most cases we can reproduce the 
  MD MSDs on long timescales within uncertainty using realizations of the 
  HDP-AR-HMM.
  \begin{itemize}
    \item We generated the MSD curves using Method 1 described in 
    Section~\ref{method:realizations} of the methods.
    %MRS: should be clearer about what ``reproduce these curves mean'', since you aren't reproducing.
    \item Our ability to reproduce these curves gives use confidence
    in their long term predictions.
    %MRS: not match under any circumstance, or without clustering (?  Not clear. 
    \item We underestimate the MSD of urea because of a small number of MD 
    trajectories with uncharacteristically high MSDs. Despite our best efforts, 
    we could not adequately parameterize an HDP-AR-HMM which reproduced said 
    behavior. (see SI)  %BJC4: TODO
    \item On long timescales, the model appear to have the same slope as 
    the MD trajectory.
  \end{itemize}

@@

  %MRS: seems odd not at least mention clustering here.

@@

  % BJC4: Not sure if this is useful.
  % MRS: it does give a good summary. But you should make it clear how you selected these 5.
  % MRS: it's implied in one of the figures, but should be stated explicitly
  By examining the parameters, it is now easier to describe the state behavior.
  \begin{enumerate}[label={State \theenumi :}, leftmargin=3.5\parindent]
     \item This state has intermediate dwell times with high past-hop dependence.
     The radial mean and evidence from Figure~\ref{fig:clustered_traj_MET} suggests
     %MRS: ``tail and pore region'' sounds the same as ``everywhere''.
     that this state is common in the tail and pore region.
     \item This state seems to occur exclusively in the tail region based on its radial
     mean. Its fluctuations are small and there is little past hop dependence
     \item This is likely a hopping state that strongly contribute to the solute's MSD. 
     It is short lived and takes very large hops. This type of behavior is exhibited 
     %MRS: ``near and far'' also sounds like everywhere 
     near and far from the pore center as discussed earlier.
     \item This is a trapping state with a strong dependence on its previous fluctuation.
     Evidence from Figure~\ref{fig:clustered_traj_MET} and a radial mean of 1.8 suggest
     that it tends to occur at intermediate distances from the pore center.
     \item This is an intermediate-length trapping state that occurs close to the pore
     center. It's fluctuations are tight and relatively independent of its previous 
     fluctuation.
  \end{enumerate}

@@

  \textit{Relating Parameters to Solute-Membrane Interactions}: Based on the
  molecules and ions which constitute the membrane, it is straightforward to 
  hypothesize two types interactions that might occur.
  \begin{itemize}  
    \item All of the solutes are polar and have hydrogen bond donating and 
    accepting groups. 
    \item The monomers have 10 oxygen atoms distributed from head to tail that are 
    capable of accepting hydrogen bonds. 
    \item Also, the ionic bond between the head groups and sodium ions can
    be easily broken since both components can be stabilized by surrounding 
    water molecules and solutes.
    \item Therefore, it is logical to search for hydrogen bonding interactions
    and solute association with sodium ions.
  \end{itemize}
  %MRS: above, start with the last point; state what the interactions are, and then give the reasoning.
  For methanol, hydrogen bonding is a much more dominant interaction than sodium
  ion association. 
  \begin{itemize}
    \item In Figure~\ref{fig:hbond_fractions}, we show that methanol participates
    in hydrogen bonds to various parts of the monomer between 27 and 45\% of the
    total time it occupies the dominant states.
    \item Methanol associates with sodium ions less than 10\% of the time in 
    dominant states.
  \end{itemize}

@@

  The local number density of heavy atoms surrounding solutes help round out methanol's
  mechanistic picture.
  \begin{itemize}
    \item Although hydrogen bonding interactions are relatively frequent, methanol molecules
    are unbound by any measurable electrostatic interactions about half or more of the 
    simulation time.  
    %MRS: this above doesn't support the thesis statement?
    \item Figure~\ref{fig:local_densities} helps to illustrate how the local density of 
    membrane components can aid in solute entrapment.
    \item When solutes experience long trapping times, the local number density of heavy 
    atoms is larger.
    \item Even when hydrogen bonds break, surrounding alkane density likely helps keep
    the solutes in place allowing them to eventually reform. This can happen deep in the
    tails or close to the monomer head groups.
    \item Methanol molecules which experience State 3 have by far the smallest local 
    number density which is likely responsible for their mobility, as suggested by the state's
    large covariance.
  \end{itemize}
